Train BrainAGE with tidymodels

### 2023-11-16 Revisions_Analyses&Figures.Rmd
Updated model prediction code for follow-up, updated readme, cleaned up analysis script.

```{r}
#troubleshooting names
aseg_labs <- brain_labels(aseg) |> 
  as.character()
aseg_vi$label[!aseg_vi$label %in% aseg_labs]
aseg_labs[!aseg_labs %in% aseg_vi$label]

@@

### Rerun old figures
```{r}
# Distributions of puberty/cognition by model
#make categorical variable for puberty
#age-corrected cognition: These are presented as Standard Scores (mean=100, SD=15).
categorical_pds_baseline_combat <-  brainage_pds_baseline_combat %>% 
  mutate(youth_pds_cat=cut(youth_mean, breaks=c(-Inf, 1, 2, 3, Inf), labels=c("1","2","3", "4")),
         parent_pds_cat=cut(parent_mean, breaks=c(-Inf, 1, 2, 3, Inf), labels=c("1","2","3", "4")),
         cognition_cat=cut(nihtbx_totalcomp_uncorrected, breaks=c(-Inf, 85, 115, Inf), labels=c(">-1", "-1-1", ">1")),
         cognition_cat_small=cut(nihtbx_totalcomp_uncorrected, breaks=c(-Inf, 92, 108, Inf), labels=c(">-.5", "-.5-.5", ">.5")),
         cog_quart = cut(nihtbx_totalcomp_uncorrected, quantile(nihtbx_totalcomp_uncorrected, na.rm=T), labels=c("0-25", "25-50", "50-75", "75-100")))
#don't z-score, include more than 1 standard deviation and less than 1 standard deviation (ignore within 1 SD)

@@

```{r}
#Youth-Report Puberty
#make categorical variable for puberty
#age-corrected cognition: These are presented as Standard Scores (mean=100, SD=15).
categorical_pds_followup_combat <-  brainage_pds_followup_combat %>% 
  mutate(youth_pds_cat=cut(youth_mean, breaks=c(-Inf, 1, 2, 3, Inf), labels=c("1","2","3", "4")),
         parent_pds_cat=cut(parent_mean, breaks=c(-Inf, 1, 2, 3, Inf), labels=c("1","2","3", "4")))

@@

```{r}
#Youth-Report Puberty
#make categorical variable for puberty
#age-corrected cognition: These are presented as Standard Scores (mean=100, SD=15).
abcd_baseline_categorical_pds_combat <- abcd_brainage_pds_baseline_combat %>% 
  mutate(youth_pds_cat=cut(youth_mean, breaks=c(-Inf, 1, 2, 3, Inf), labels=c("1","2","3", "4")),
         parent_pds_cat=cut(parent_mean, breaks=c(-Inf, 1, 2, 3, Inf), labels=c("1","2","3", "4")),
         cognition_cat=cut(nihtbx_totalcomp_uncorrected, breaks=c(-Inf, 85, 115, Inf), labels=c(">-1", "-1-1", ">1")),
         cognition_cat_small=cut(nihtbx_totalcomp_uncorrected, breaks=c(-Inf, 92, 108, Inf), labels=c(">-.5", "-.5-.5", ">.5")),
         cog_quart = cut(nihtbx_totalcomp_uncorrected, quantile(nihtbx_totalcomp_uncorrected, na.rm=T), labels=c("0-25", "25-50", "50-75", "75-100")))

@@

```{r}

#Youth-Report Puberty
#make categorical variable for puberty
#age-corrected cognition: These are presented as Standard Scores (mean=100, SD=15).
abcd_categorical_pds_followup_combat <-  abcd_brainage_pds_followup_combat %>% 
  mutate(youth_pds_cat=cut(youth_mean, breaks=c(-Inf, 1, 2, 3, Inf), labels=c("1","2","3", "4")),
         parent_pds_cat=cut(parent_mean, breaks=c(-Inf, 1, 2, 3, Inf), labels=c("1","2","3", "4")))

@@

```{r}
# Model feature - follow-up

@@

figure out how many baseline training v analysis are in follow-up
```{r}
# check how many from training sample are in follow-up

@@

Plot of longitudinal BrainAGE
```{r}
# Spaghetti plot, show direction of gap change
# is x-axis time, or age?
# this might just be too many points to show? maybe fit a line over the top
# vlad's model
brainage_both_waves_long <- rbind(brainage_pds_baseline_combat %>% select(src_subject_id, truth, corrected_pred, corrected_gap) %>% mutate(wave=1), brainage_pds_followup_combat %>% select(src_subject_id, truth, corrected_pred, corrected_gap)%>% mutate(wave=2))

@@

# could also facet by slope to reduce lines

# abcd model
brainage_abcd_both_waves_long <- rbind(abcd_brainage_pds_baseline_combat %>% select(src_subject_id, truth, corrected_pred, corrected_gap) %>% mutate(wave=1), abcd_brainage_pds_followup_combat %>% select(src_subject_id, truth, corrected_pred, corrected_gap)%>% mutate(wave=2))

>>>>

Optional feature name troubleshooting for ggseg
```{r}
#troubleshooting names
aseg_labs <- brain_labels(aseg) |> 
  as.character()
aseg_vi$label[!aseg_vi$label %in% aseg_labs]
aseg_labs[!aseg_labs %in% aseg_vi$label]

@@

### Rerun old figures
# Distributions - Baseline (Model 1)
```{r}
# Distributions of puberty/cognition by model
#make categorical variable for puberty
#age-corrected cognition: These are presented as Standard Scores (mean=100, SD=15).
categorical_pds_baseline_combat <-  brainage_pds_baseline_combat %>% 
  mutate(youth_pds_cat=cut(youth_mean, breaks=c(-Inf, 1, 2, 3, Inf), labels=c("1","2","3", "4")),
         parent_pds_cat=cut(parent_mean, breaks=c(-Inf, 1, 2, 3, Inf), labels=c("1","2","3", "4")),
         cognition_cat=cut(nihtbx_totalcomp_uncorrected, breaks=c(-Inf, 85, 115, Inf), labels=c(">-1", "-1-1", ">1")),
         cognition_cat_small=cut(nihtbx_totalcomp_uncorrected, breaks=c(-Inf, 92, 108, Inf), labels=c(">-.5", "-.5-.5", ">.5")),
         cog_quart = cut(nihtbx_totalcomp_uncorrected, quantile(nihtbx_totalcomp_uncorrected, na.rm=T), labels=c("0-25", "25-50", "50-75", "75-100")))
#don't z-score, include more than 1 standard deviation and less than 1 standard deviation (ignore within 1 SD)

@@

# Distributions - Follow-Up (Model 1)
```{r}
#Youth-Report Puberty
#make categorical variable for puberty
#age-corrected cognition: These are presented as Standard Scores (mean=100, SD=15).
categorical_pds_followup_combat <-  brainage_pds_followup_combat %>% 
  mutate(youth_pds_cat=cut(youth_mean, breaks=c(-Inf, 1, 2, 3, Inf), labels=c("1","2","3", "4")),
         parent_pds_cat=cut(parent_mean, breaks=c(-Inf, 1, 2, 3, Inf), labels=c("1","2","3", "4")))

@@

# Distributions - Baseline (Model 2)
```{r}
#Youth-Report Puberty
#make categorical variable for puberty
#age-corrected cognition: These are presented as Standard Scores (mean=100, SD=15).
abcd_baseline_categorical_pds_combat <- abcd_brainage_pds_baseline_combat %>% 
  mutate(youth_pds_cat=cut(youth_mean, breaks=c(-Inf, 1, 2, 3, Inf), labels=c("1","2","3", "4")),
         parent_pds_cat=cut(parent_mean, breaks=c(-Inf, 1, 2, 3, Inf), labels=c("1","2","3", "4")),
         cognition_cat=cut(nihtbx_totalcomp_uncorrected, breaks=c(-Inf, 85, 115, Inf), labels=c(">-1", "-1-1", ">1")),
         cognition_cat_small=cut(nihtbx_totalcomp_uncorrected, breaks=c(-Inf, 92, 108, Inf), labels=c(">-.5", "-.5-.5", ">.5")),
         cog_quart = cut(nihtbx_totalcomp_uncorrected, quantile(nihtbx_totalcomp_uncorrected, na.rm=T), labels=c("0-25", "25-50", "50-75", "75-100")))

@@

# Distributions - Follow-Up (Model 2)
```{r}
#Youth-Report Puberty
#make categorical variable for puberty
#age-corrected cognition: These are presented as Standard Scores (mean=100, SD=15).
abcd_categorical_pds_followup_combat <-  abcd_brainage_pds_followup_combat %>% 
  mutate(youth_pds_cat=cut(youth_mean, breaks=c(-Inf, 1, 2, 3, Inf), labels=c("1","2","3", "4")),
         parent_pds_cat=cut(parent_mean, breaks=c(-Inf, 1, 2, 3, Inf), labels=c("1","2","3", "4")))

@@

# Model features - follow-up
```{r}

@@

Check how many baseline training v analysis are in follow-up
```{r}
# check how many from training sample are in follow-up

@@

Plot of longitudinal BrainAGE
```{r}
# Spaghetti plot, show direction of gap change

# Existing model
brainage_both_waves_long <- rbind(brainage_pds_baseline_combat %>% select(src_subject_id, truth, corrected_pred, corrected_gap) %>% mutate(wave=1), brainage_pds_followup_combat %>% select(src_subject_id, truth, corrected_pred, corrected_gap)%>% mutate(wave=2))

@@

# ABCD model
brainage_abcd_both_waves_long <- rbind(abcd_brainage_pds_baseline_combat %>% select(src_subject_id, truth, corrected_pred, corrected_gap) %>% mutate(wave=1), abcd_brainage_pds_followup_combat %>% select(src_subject_id, truth, corrected_pred, corrected_gap)%>% mutate(wave=2))

### 2023-11-16 UpdatedModels/ABCDModelTraining_FollowUp.Rmd
Updated model prediction code for follow-up, updated readme, cleaned up analysis script.

@@

#workflow_predict <- predict(fit_workflow, analysis_sample)

>>>>

Alternate/Backup Prediction - Use if prior version doesn't work
```{r}
# The following version converts the workflow object to a raw xgb object. Over the course of the project, we've discovered that certain features in the xgboost package haven't always worked with updates within tidymodels. In case of deprecation of certain functions/ (or issues with version compatibility), this code should let you access the actual xgboost object, which can then be used to make model predictions. I don't anticipate this will be needed, but if the code on lines 221-233 doesn't work, this might!

abcd_model_obj_combat_followup <- fit_workflow_combat_followup$fit$fit$fit

@@
